% \begin{acknowledgements}
\section*{Acknowledgements}
First and foremost, we wish to recognize and acknowledge the significant cultural role and reverence that the summit of Maunakea has always had within the Indigenous Hawaiian community. We are grateful and thank the community for the privilege of building instrumentation on and conducting observations from this sacred Mauna.

We thank Joshua Hacko (NH Micro), Iven Hamilton (UH), Shane Jacobson (UH/Gemini), Michael Lemmin (NAOJ), Paul Sumner (Opto-line), Lucio Ramos (NAOJ), Paul Toyoma (UH), and Matthew Wung (NAOJ) for their time and expertise in the manufacturing, assembly, testing, and integration of the VAMPIRES upgrades.

We thank Javier Alcoloa for the valuable discussions on our R Aqr observations. We thank Timothy Brandt and collaborators for the use of data obtained under Subaru proposal ID o23149. We thank Michael Fitzgerald and collaborators for the use of data obtained under Subaru proposal ID o23150. We thank telescope operators Sean Cunningham, Erin Dailey, Jonathan Laverty, Michael Letawsky, Andrew Neugarten, Larissa Schumacher, and Aidan Walk for their support during the observations used in this work.

This research was funded by the Heising-Simons Foundation through grant \#2020-1823. Based on data collected at Subaru Telescope, which is operated by the National Astronomical Observatory of Japan. The development of SCExAO is supported by the Japan Society for the Promotion of Science (Grant-in-Aid for Research \#23340051, \#26220704, \#23103002, \#19H00703, \#19H00695, and \#21H04998), the Subaru Telescope, the National Astronomical Observatory of Japan, the Astrobiology Center of the National Institutes of Natural Sciences, Japan, the Mt Cuba Foundation and the Heising-Simons Foundation.
% \end{acknowledgements}